One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction

### 2017-11-30 draft_04.tex
edited first (preparatory) section

\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \usepackage{xcolor}
 \usepackage{bm}
 \usepackage[]{hyperref}
\hypersetup{
    colorlinks = false,
    linkbordercolor = {red},
    citebordercolor = {green},
}
\usepackage{epstopdf}
 \begin{document}
 \title{One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction}
 \author{Zhiyu}
 \affiliation{Fudan University, Shanghai, China}

@@

@@

On the other hand, in a nonquilibrium many-body system, a more intriguing question is about its
thermalization or relaxation property. Thermalization is among the most fundamental processes for a
many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam
paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a
long time. A lot of research on thermalization property has been done in other classical
system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with
Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment
by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's
thermalization time scales from the perspective of canonical distribution as well as Lyapunov
exponent. %With these approaches, we will show that the thermalization time scale manifest a
          %competition between total energy and interaction energy, so there is not any singular
          %behavior in the thermalization time scales except the trivial divergence at zero
          %interaction.

\section{Preparation}\label{section:preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F_0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}

Using the transformation Eq.\ref{eq:transform}, we can rewrite Eq.\ref{eq:preparation2} into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\label{eq:transform}
\end{equation}

In this manner, we reduced the number of the parameters in our model to three: $\tilde{H}$,
$\tilde{F_0}$, $N$. Equivalently, we set $\sigma$, $\omega_0$ and $m$ to 1 in our numerical
simulation. Throughout, we will use ``E" and ``$F_0$" to denote their reduced version $\tilde{H}$,
$\tilde{F_0}$.

\subsection{Overview}
\subsubsection{Relation between Radius and Energy}
The radius of the cloud R is defined as the root-mean-square of particles' position $\left\lbrace x_i\right\rbrace$
\begin{equation}
R\equiv\left(\overline{x_i^2}\right)^\frac{1}{2}
\label{eq:def_of_R}
\end{equation}. When E is very large, the interaction energy can be omitted, so E is dominated by the kinetic energy and potential energy of harmonic oscillators. Thus, $E=\omega_0^2\sum_{i}x_i^2=N\omega_0^2R^2$

\subsubsection{Ground states}

For the ground state of this system, one could make the following prediction: \\(1) When the
$\sigma$ is larger than the size of the cloud (in view of the scaling behavior, large $\sigma$ is
equivalent to small $F_0$ limit when we fix $\sigma=1$), it can be easily shown that the ground
state is a solid-like state, where each particle keeps an equal distance ($\frac{2F_0}{\omega_0^2}$)
with its nearest neighbors . The low-lying excitation is particles' independent harmonic oscillation
around their equilibrium position.\\(2) When $F_0$ is very large the molecules act like a hardcore
gas, while the size of particle is just $\sigma$.

\begin{figure}[h]
\centering
\includegraphics[scale=0.6]{ZhiyuPictures/N=5_GS_pre_2_rev.eps}
\caption{particles' position at ground states }
\label{fig:GS1}
\end{figure}

@@

@@

\begin{figure}[h
]
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/bounce.png}
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/pass.png}
\end{minipage}
}

\caption{a schematic diagram of the collision process: \textbf{(a)}When $F_0\sigma$ is large, particles can not pass each other. Subduing to a constant repulsive force in collision, their trajectory in rotating frame is a circular arc.  \textbf{(b)} When $F_0\sigma$ is small, particles easily pass each other, so that the force they feel is not continuous. Their trajectory has a turning point. The dashed bright blue trajectory is identical with the one in (a), which is the critical situation between passing and bouncing. In that case, particles have zero relative velocity when they collide. }
\label{fig:Breathingfrequency3}
\end{figure}

@@

\caption{the dependence of beat frequency on internal energy. The diagram is the energy of two particles. Here, $F_0$ is set to be 1. In this case, the larger the $E_i$, the lower the beat frequency. \label{fig:thermalization1}}%Actually in very low energy, we can retrieve a beat, but that is for different reason -- it is a beat between $\omega_0$ and $\frac{\omega_0}{2}+\delta $

>>>>

\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \usepackage{xcolor}
 \usepackage{bm}
 \usepackage[]{hyperref}
 \hypersetup{
    colorlinks = false,
    linkbordercolor = {red},
    citebordercolor = {green},
}

 \usepackage{epstopdf}

\begin{document}
 \title{One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction}
 \author{Zhiyu}
 \affiliation{Fudan University, Shanghai, China}

\date{\today}

@@

Will write Introduction after everything else is written.

---------

Stat-mech from the `bottom up'.  Find papers that do the same.

More detailed citations of breathing mode.  There are also papers on classical breathing mode, e.g.,
by Bonitz.

@@

On the other hand, in a nonquilibrium many-body system, a more intriguing question is about its
thermalization or relaxation property. Thermalization is among the most fundamental processes for a
many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam
paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a
long time. A lot of research on thermalization property has been done in other classical
system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with
Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment
by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's
thermalization time scales from the perspective of canonical distribution as well as Lyapunov
exponent.

%With these approaches, we will show that the thermalization time scale manifest a
%competition between total energy and interaction energy, so there is not any singular
%behavior in the thermalization time scales except the trivial divergence at zero
%interaction.

------------------

Breathing frequency

@@

------------------

\section{The model and its equilibrium properties}\label{section:preparation}

\subsection{Model and Scaling}

Our model for the interacting classical gas involves particles with a simple finite-range repulsive
interaction.  Two particles repel each other whenever they are within a distance $\sigma$ from each
other.  The force of repulsion is a constant, $F_0$, within this distance and zero when the distance
is larger. 
%
\begin{equation}
F(|x|) ~=~ \begin{cases}
F_0  &\quad |x|<\sigma\\
0  &\quad |x|>\sigma
\end{cases}
\label{eq:preparation1}
\end{equation}
where $|x|$ is the distance between two particles.  This equation describes the magnitude of the
interparticle force; the direction is always repulsive.

The gas contains $N$ such identical particles, each of mass $m$, in a harmonic trap.  The
Hamiltonian describing the gas is 
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i
v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
%
It is useful to rescale the quantities, and measure distance, time, energy and force in units of
$\sigma$, $1/\omega$, $m\omega_0^2\sigma^2$ and $m\omega_0^2\sigma$ respectively:
\begin{equation} \label{eq:transform}
  \tilde{x_i} = \frac{x_i}{\sigma}\ , \quad 
  \tilde{H} = \frac{H}{m\omega_0^2\sigma^2}\ ,  \quad \tilde{t} = \omega_0t \ ,  \quad 
\tilde{F_0} = \frac{F_0}{m\omega_0^2\sigma}\ .
\end{equation}
%
Eq.~\ref{eq:preparation2} is then rewritten as
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
%
Through this rescaling, we have reduced the number of the parameters in our model to three: the
energy $\tilde{H}$, the interaction strength $\tilde{F_0}$, and the number of particles $N$.  The
rescaling is equivalent to setting $\sigma$, $\omega_0$ and $m$ to $1$; this is what we do in our
numerical simulation.  In the rest of this paper, we will use ``E" and ``$F_0$" to denote the
reduced versions $\tilde{H}$ and $\tilde{F_0}$.

We will be concerned with the size of the cloud, which we quantify through the root-mean-square of
particle positions  $\left\lbrace x_i\right\rbrace$
\begin{equation}
R\equiv\left(\overline{x_i^2}\right)^\frac{1}{2}
\label{eq:def_of_R}
\end{equation}
and refer to as the `radius' or `cloud radius'.

In the parameter regime where the energy $E$ is very large ($E\gg F_0$), it is dominated by the trap
energy, i.e., the interaction energy is negligible.  Thus,
$E\approx\omega_0^2\sum_{i}x_i^2=N\omega_0^2R^2$ in this regime

\begin{figure}[tbhp]
\centering
\includegraphics[scale=0.6]{ZhiyuPictures/N=5_GS_pre_2_rev.eps}
\caption{The ground state of a system with $N=5$ particles, shown via the positions of the
  particles.  At large interactions, the particles position themselves just outside the range of
  interactions of the neighboring particles.}
\label{fig:GS1}
\end{figure}

\subsection{Ground states}

The lowest-energy state of the system has zero kinetic energy; in this state the particles find
stationary positions which minimize the trap (potential) and interaction energies.  The trap
potential tries to squeeze the particles at the trap center, while the interaction tries to push
them apart.

Because of the `Heaviside theta' form of our interaction, at large enough interactions (large $F_0$)
the particles are spaced exactly at distance equal to the range $\sigma$, which is distance $1$ in
our rescaled units.  The interaction is effectively a `hard-core' interaction.

At smaller $F_0$, the energy is minimized by having the particles approximately equidistant, at
distance $\approx 2F_0/\omega_0^2$ from each other.  In this `solid'-like state, the low-lying
excitation involves independent oscillation of the particles around their equilibrium position.

@@

\begin{figure}[h
]
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/bounce.png}
\end{minipage}
}
\subfigure[]{
\begin{minipage}[b]{0.4\textwidth}
\includegraphics[width=1\textwidth]{ZhiyuPictures/pass.png}
\end{minipage}
}
\caption{a schematic diagram of the collision process: \textbf{(a)}When $F_0\sigma$ is large, particles can not pass each other. Subduing to a constant repulsive force in collision, their trajectory in rotating frame is a circular arc.  \textbf{(b)} When $F_0\sigma$ is small, particles easily pass each other, so that the force they feel is not continuous. Their trajectory has a turning point. The dashed bright blue trajectory is identical with the one in (a), which is the critical situation between passing and bouncing. In that case, particles have zero relative velocity when they collide. }
\label{fig:Breathingfrequency3}
\end{figure}

@@

\caption{the dependence of beat frequency on internal energy. The diagram is the energy of two
  particles. Here, $F_0$ is set to be 1. In this case, the larger the $E_i$, the lower the beat
  frequency. \label{fig:thermalization1}}
%Actually in very low energy, we can retrieve a beat, but that is for different reason -- it is a
%beat between $\omega_0$ and $\frac{\omega_0}{2}+\delta $
